---
output: 
  html_document:
    css: assessment_platform.css

params:
  compartment: must_supply
  series: must_supply
---

```{r pagetitle, include = FALSE}
page_title <- params$series
page_title <- gsub("_", " ", page_title)
page_title <- gsub(" Not applicable", "", page_title)
page_title <- stringi::stri_trans_general(page_title, "LATIN-ASCII")

```

---
pagetitle: `r page_title`
---


```{r load_packages, include = FALSE}
require(lattice)
require(grid)
require(knitr)


# will also require libraries stringi and stringr to be installed, but not loaded

knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
options(width = 110)
```


```{r setup_data, echo = FALSE}

#### get key structures ####

assessment_object <- get(paste0(params$compartment, "_web"))

determinands <- assessment_object[["determinands"]]

assessment_object <- assessment_object[["assessment"]]


##### get data and assessments in the station / determinand group ####

# biota: need further subdivision by species and matrix (contaminants only)
# convert detGroup to character just to be on safe side!

assessment_object$timeSeries <- within(assessment_object$timeSeries, {
  detGroup <- as.character(detGroup)
})

var_id <- c("station", "detGroup")
if (params$compartment == "biota") {
  var_id <- c(var_id, "species")
  group_id <- assessment_object$timeSeries[params$series, "detGroup"]
  if (!group_id  %in% c("Effects", "Imposex", "Metabolites"))
    var_id <- c(var_id, "matrix")
  species_id <- assessment_object$timeSeries[params$series, "species"]
  if (get.info("species", species_id, "family") %in% "Mammal")
    var_id <- c(var_id, "AMAP_group")
}

assessment_object <- within(assessment_object, {
  timeSeries$multi_series <- do.call("paste", timeSeries[var_id])
})

id <- assessment_object$timeSeries[params$series, "multi_series"]
assessment_object <- ctsm.subset.assessment(assessment_object, multi_series %in% id)

# sort, using determinands, to ensure everything is in the 'correct' presentational order

assessment_object <- within(assessment_object, {
  timeSeries <- timeSeries[order(match(timeSeries$determinand, determinands)), ]
})


#### get data, assessment and info for params$series ####

ok <- with(assessment_object$data, seriesID %in% params$series)
data <- assessment_object$data[ok, ]

assessment <- assessment_object$assessment[[params$series]]

info <- with(assessment_object, c(info, timeSeries[params$series, ]))

# determinand info

info$group <- unique(as.character(data$group))

info$det_name <- get.info("determinand", info$determinand, "common.name")
if (info$group %in% c("Metals", "PAH_parent", "Organotins")) {
  info$det_name <- tolower(info$det_name)
}

info$good.status <- get.info("determinand", info$determinand, "good.status")
info$good.status <- as.character(info$good.status)

info$distribution <- get.info("determinand", info$determinand, "distribution")
info$distribution <- as.character(info$distribution)


# species info

if (info$compartment %in% "biota") {
  info$species_name <- get.info("species", info$species, "common.name")
  ok <- sapply(
    c("European", "Manila", "Atlantic", "Mediteranean", "Baltic", "Eurasian"), 
    grepl, 
    info$species_name
  )
  if (!any(ok)) info$species_name <- tolower(info$species_name)
  
  if (info$species == "Hinia reticulata") {
    info$species_name <- "dog whelk"
  }
  
  info$species_group <- get.info("species", info$species, "family")
  info$species_group <- as.character(info$species_group)
}

if (info$compartment %in% "biota")
  info$AMAP_group <- gsub("_", " ", info$AMAP_group)


if (info$group %in% c("Effects", "Imposex")) {
  txt_concentration <- "level"
} else {
  txt_concentration <- "concentration"
}

txt_concentrations <- paste0(txt_concentration, "s")

txt_measurements <- switch(
  txt_concentration, 
  concentration = "concentration measurements",
  "measurements"
)
  
txt_compounds <- if (info$group %in% c("Imposex", "Effects", "Metabolites")) {
  "responses"
} else {
  "compounds"
}


# station info

if (is.na(info$stationName)) info$stationName <- info$station

# matrix info

# can be multiple matrices for sediment, in which case take most common
# should only be a single matrix for other compartments

info$matrix <- unique(as.character(data$matrix))

if (params$compartment != "sediment" & length(info$matrix) > 1) 
  cat("Error: multiple matrices - need to investigate")

if (length(info$matrix > 1)) {
  wk <- table(as.character(data$matrix))
  info$matrix <- names(wk[which.max(wk)])
}
 
info$matrix_name <- get.info("matrix", info$matrix, "name")

info$matrix_name <- switch(
  info$matrix_name, 
  "erythrocytes (red blood cells in vertebrates)" = "red blood cells",
  "egg homogenate of yolk and albumin" = "egg yolk and albumin",
  "hair/fur" = "hair",
  info$matrix_name
)



# sex info

if (info$compartment == "biota")
  info$sex_name <- switch(as.character(info$sex), M = "male", F = "female")


# units 

info$unit <- as.character(get.info("determinand", info$determinand, "unit", info$compartment))

info$basis <- as.character(info$basis)

args.list <- list(units = info$unit, basis = info$basis, html = TRUE, compartment = info$compartment)

is_normalised <- info$compartment == "sediment" | info$basis %in% "L"

if (info$compartment == "sediment") {
  extra.text <- if (info$determinand %in% c("CD", "PB")) "5% aluminium" else "5% organic carbon"
  extra.text <- paste("normalised to", extra.text)
  args.list <- c(args.list, extra.text)
}  

if (info$basis %in% "L") {
  extra.text <- "normalised to 5% lipid"
  args.list <- c(args.list, extra.text)
}  

info$unit_text <- do.call(label.units, args.list)


# assessment concentrations

is_AC <- any(!is.na(assessment$AC))

if (is_AC) {
  assessment$AC <- na.omit(assessment$AC)
  
  AC_id <- names(assessment$AC)
  
  AC_type <- dplyr::case_when(
    AC_id %in% c("BAC", "NRC") ~ "BAC",
    AC_id %in% c("MPC", "QShh") ~ "HQS",
    TRUE ~ "EAC"
  )

  if (max(table(AC_type)) > 1L) {
    stop("multiple BACs, EACs or HQSs identified")
  }
  
  is_BAC <- "BAC" %in% AC_type
  if (is_BAC) {
    BAC_id <- AC_id[match("BAC", AC_type)]
  }

  is_EAC <- "EAC" %in% AC_type
  if (is_EAC) {
    EAC_id <- AC_id[match("EAC", AC_type)]
  }

  is_HQS <- "HQS" %in% AC_type
  if (is_HQS) {
    HQS_id <- AC_id[match("HQS", AC_type)]
  }

  
  # remove na from reference.values - needed to be compatible with AC_id
    
  assessment$reference.values <- na.omit(assessment$reference.values)
}


# is there a non-parametric test of status

status_nonparametric <- 
  if (info$group %in% "Imposex") {
    FALSE
  } else {
    is_AC && !is.na(assessment$summary[[paste0(AC_id[1], "below")]])
  }

status_nonparametric <- FALSE


# is there a parametric model fit

is_pred <- "pred" %in% names(assessment)


# trend information 

anova <- assessment$anova
coefficients <- assessment$coefficients

# deal with full model in some survival models - need to resolve

if (info$distribution == "survival" && !is.null(anova) && "full" %in% row.names(anova)) {
  pos <- match("full", row.names(anova))
  anova <- anova[-pos, ]
}

anova_ok <- if (info$detGroup == "Imposex") {
  !is.null(anova) || !is.null(coefficients)  
} else {
  !is.null(anova) && nrow(anova) > 1
}

# is there a change point model 
# default is FALSE, but can be overridden in imposex assessments

is_change <- FALSE


#### get data, assessment and info for multiplots ####

info_multi <- info

info_multi <- within(info_multi, {
  seriesID <- row.names(assessment_object$timeSeries)
  determinand <- as.character(assessment_object$timeSeries$determinand)
  names(determinand) <- seriesID
})

# get series names for labelling plots - usually just determinand, but could be more complex if e.g.
# measured in multiple tissues, or when dealing with biological effects

info_multi$plotNames <- with(info_multi, list(data = determinand, assessment = determinand))

if (any(duplicated(info_multi$plotNames))) {
  dups <- with(info_multi, duplicated(determinand) | duplicated(determinand, fromLast = TRUE))
  if (info$compartment %in% c("biota", "sediment")) {
    info_multi$plotNames$data[dups] <- paste(
      info_multi$determinand, assessment_object$timeSeries$level6name, sep = "\n")[dups]
    info_multi$plotNames$assessment[dups] <- paste(
      info_multi$determinand, assessment_object$timeSeries$level6name)[dups]
  } else {
    wk <- assessment_object$timeSeries$filtered
    wk <- ifelse(wk == "Yes", "filtered", "unfiltered")
    info_multi$plotNames$data[dups] <- paste(info_multi$determinand, wk, sep = "\n")[dups]
    info_multi$plotNames$assessment[dups] <- paste(info_multi$determinand, wk)[dups]
  }
  if (any(duplicated(info_multi$plotNames$data))) 
    cat("Error: duplicate plotting names - need to extend coding")
}
  
info_multi$matrix <- with(assessment_object$data, tapply(matrix, seriesID, function(x) {
  out <- table(as.character(x))
  names(out[which.max(out)])
}))

if (info$compartment == "water") {
  info_multi$filtered <- with(assessment_object$data, tapply(filtered, seriesID, function(x) {
    out <- table(as.character(x))
    names(out[which.max(out)])
  }))
}
info_multi$group <- info$group

is_multi <- length(info_multi$plotNames$data) > 1L 

is_ratio_dets <- function(det1, det2) all(c(det1, det2) %in% info_multi$determinand)

is_ratio <- switch(
  info$group, 
  Metals = info$compartment %in% "biota" && 
    info$determinand %in% c("HG", "SE") &&
    is_ratio_dets("HG", "SE"),
  PAH_parent = info$compartment %in% c("biota", "sediment") &&
    info$determinand %in% c("ANT", "PA", "FLU", "PYR", "ICDP", "BGHIP", "BAA", "CHR") &&
    (is_ratio_dets("ANT", "PA") | is_ratio_dets("FLU", "PYR") | 
       is_ratio_dets("ICDP", "BGHIP") | is_ratio_dets("BAA", "CHR")),
  PBDEs = info$compartment %in% "biota" &&
    info$determinand %in% c("BDE47", "BD153") && 
    is_ratio_dets("BDE47", "BD153"),
  Organofluorines = info$compartment %in% "biota" &&
    info$determinand %in% c("PFNA", "PFOA", "PFUNDA", "PFDA", "PFTRDA", "PFDOA") &&
    (is_ratio_dets("PFNA", "PFOA") | is_ratio_dets("PFUNDA", "PFDA") | 
       is_ratio_dets("PFTRDA", "PFDOA")),
  Organochlorines = 
    ((info$compartment %in% "biota" && info$species_group %in% "Bivalvia") | 
       (info$compartment %in% "sediment")) && 
    info$determinand %in% c("DDEPP", "DDTPP", "DDTOP") && 
    (is_ratio_dets("DDEPP", "DDTPP") | is_ratio_dets("DDTOP", "DDTPP")),
  FALSE
)

is_ratio <- FALSE

```


```{r intro_txt, include = FALSE}
wk <- if (info$species != "Tritia nitida / reticulata") info$species_name else "dog whelk"
intro_txt <- switch(
  info$compartment, 
  biota = paste(wk, info$matrix_name), 
  info$compartment
)

```

### Assessment plots and statistical analysis {.tabset}

This report provides details of the assessment of `r paste(info$det_name, txt_concentrations)` in `r intro_txt` at station `r info$stationName`. `r if (is_normalised) paste0("Concentrations are ", extra.text, " unless otherwise stated. ")`

<ul>
<li class = "gap">Timeseries metadata.  Key information about the timeseries.</li>

```{r txt_variables, include = FALSE}
txt_trend1 <- paste(
  "The",
  if (is_pred) "trend is" else "points are", 
  "plotted on the log scale, with the axis labels back-transformed for ease of interpretation."
)

txt_trend2 <- switch(
  info$distribution,
  lognormal = paste0(
    "The points are the annual medians of the individual log ", txt_measurements, 
    ", with solid circles denoting uncensored values and < denoting less-than values."
  ),
  "The points are the annual means of the individual measurements with pointwise 90% confidence limits."
)
  
txt_trend3 <- switch(
  info$distribution, 
  lognormal = paste0(
    "The points are the individual ", txt_measurements, 
    ", with + denoting uncensored values, ", 
    "d or q denoting values below the limit of detection or quantification and ",
    "< denoting other less-than values."
  ), 
  "The points are the individual measurements."
)

txt_trend4 <- switch(
  info$distribution, 
  lognormal = paste(
    "There are too few years, or too many less-than values, to fit a model to these data.", 
    "Instead, the solid line shows an estimate of the mean concentration in the period 2016-2021", 
    "and the grey shaded areas are pointwise 90% confidence bands based on estimates of", 
    "variability from timeseries at other monitoring stations. "
  ), 
  "There are too few years to fit a model to these data. "
)

txt_trend5 <- switch(
  info$distribution, 
  lognormal = paste(
    "The solid line shows an estimate of the mean concentration in the period", 
    "2016-2021 and the grey shaded areas are pointwise 90% confidence bands", 
    "based on estimates of variability from timeseries at other monitoring stations."
  ),
  paste(
    "The solid line shows an estimate of the mean level in the last monitoring year",
    "and the grey shaded areas are pointwise 90% confidence bands based on the",
    "individual measurements that year."
  )
)




txt_AC <- paste(
  "The threshold value is", 
  "indicated on the right hand side of the plot and by a horizontal dashed line"
)

txt_AC <- paste0(
  txt_AC, 
  switch(
    info$distribution, 
    lognormal = " (if within range of the data).",
    "."
  )
)
```

```{asis, eval = is_pred}
<li class = "gap">Assessment plot.  The fitted trend (solid line) with pointwise 90% confidence bands (grey shaded areas).  `r if (info$distribution == "lognormal") txt_trend1` `r txt_trend2` `r if (is_AC) txt_AC`</li>
<li class = "gap">Trend with data.  The fitted trend with pointwise 90% confidence bands.  `r txt_trend3`</li>
```  

```{asis, eval = !is_pred}
<li class = "gap">Assessment plot. `r txt_trend2` `r if (info$distribution == "lognormal")  txt_trend1` `r txt_trend4``r if (is_AC) txt_AC`</li>
<li class = "gap">Trend with data.  `r txt_trend3` `r txt_trend5`</li>
```  


```{asis, eval = info$compartment %in% "biota"}
<li class = "gap">Auxiliary data.  The individual `r txt_measurements` with supporting information: mean length of the individuals in the sample; dry weight (%) of the sample; lipid weight (%) of the sample.</li>
```

```{asis, eval = info$compartment %in% "sediment"}
<li class = "gap">Auxiliary data.  The individual concentration measurements`r if (is_normalised) paste0(", both non-normalised and ", extra.text, ",")` with supporting information: `r if (info$determinand %in% c("CD", "PB")) "aluminium content (%) of the sample; lithium content (%) of the sample." else "organic carbon content (%) of the sample; loss on ignition (%) of the sample."`</li>
```

```{asis, eval = info$compartment %in% "water"}
<li class = "gap">Auxiliary data.  No auxiliary data are currently plotted.</li>
```


```{r txt_related_compounds, include = FALSE}
txt_related1 <- switch(
  info$group, 
  Imposex = "There are no related responses.",
  Effects = paste(
    "Assessment plots for all the biological effects (other than PAH metabolites) measured in", 
    info$species_name, "at this station."
  ),
  Metabolites = paste(
    "Assessment plots for all the PAH metabolites measured in", 
    info$species_name, "at this station."
  ),
  paste0(
    "Assessment plots for all the compounds in the same chemical group measured ", 
    if (info$compartment == "biota") paste0("in ", info$species_name, " ", info$matrix_name, " "),
    "at this station."
  )
)

txt_related2 <- switch(
  info$group, 
  Imposex = "There are no related responses.",
  Effects = paste(
    "A scatter plot matrix of all the individual biological effects measurements", 
    "(other than PAH metabolites) in", 
    info$species_name, "at this station."
  ),
  Metabolites = paste(
    "A scatter plot matrix of all the individual PAH metabolite measurements in", 
    info$species_name, "at this station."
  ),
  paste0(
    "A scatter plot matrix of all the individual concentration measurements for all the ",
    "compounds in the same chemical group ",
    if (info$compartment == "biota") paste0("in ", info$species_name, " ", info$matrix_name, " "),
    "at this station."
  )
)
```

<li class = "gap">Assessment (related `r txt_compounds`).  `r txt_related1`</li>
<li class = "gap">Data (related `r txt_compounds`). `r txt_related2`</li>



```{r, include = FALSE}
txt <- if (anova_ok) "Finally, the tab" else "The tab also"

txt_stat_1 <- switch(
  info$distribution, 
  lognormal = "concentration in the period 2016-2021",
  "level in the last monitoring year"
)

```

```{asis, eval = is_pred}
<li class = "gap">Statistical analysis. This tab summarises the fits of models with different amount of smoothing. The final choice is the model with the minimum `r if ("AICc" %in% names(anova)) "AICc" else "AIC"` (with the caveat that a linear model is always preferred to a mean model if there are sufficient years of data). `r if (anova_ok) "The tab also gives the contrast between the fitted values at the start and end of the timeseries, and over the most recent twenty years. "` `r if (is_AC) paste(txt, "assesses status by comparing the fitted value at the end of the timeseries to the threshold value.")`</li>
```

```{asis, eval = !is_pred & !status_nonparametric}
<li class = "gap">Statistical analysis. This tab assesses status by comparing the upper confidence interval on the mean `r txt_stat_1`  to the threshold value.</li>
```

```{asis, eval = !is_pred & status_nonparametric}
<li class = "gap">Statistical analysis. This tab gives the results of a non-parametric status assessment.</li>
```


</ul>


<br>

#### Timeseries metadata

<br>

<ul>
<li>Subbasin: `r info$region`</li>
<li>L3 area: `r info$l3area`</li>
<li>L4 area: `r info$l4area`</li>
<li>Station code: `r info$station`</li>
<li>Station name: `r info$stationName`</li>
<li>Station latitude: `r format(info$latitude, digits = 2, nsmall = 2)`</li>
<li class = "gap">Station longitude: `r format(info$longitude, digits = 2, nsmall = 2)`</li>
<li>Compartment: `r info$compartment`</li>

```{asis, eval = info$compartment %in% "biota"}
<li>Species latin name: `r if (info$species != "Tritia nitida / reticulata") info$species else "Hinia reticulata"`</li>
<li>Species common name: `r if (info$species != "Tritia nitida / reticulata") info$species_name else "Dog whelk"`</li>  
```


<li class = "gap">`r switch(info$compartment, biota = "Tissue", "Matrix")`: `r info$matrix_name`</li>
<li>Determinand code: `r info$determinand`</li>
<li>Determinand name: `r info$det_name`</li>

```{asis, eval = info$detGroup %in% "PAH metabolites"}
<li>Method of chemical analysis: `r info$metoa`</li>
```

<li class = "gap">Units: `r info$unit_text`</li>
<li>Extraction: `r info$extraction`</li>
</ul>

<br>




#### Assessment plot

```{r assessment_plot, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.data(data, assessment, info, type = "assessment", xykey.cex = 1.4)
```


#### Trend with data

```{r data_plot, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.data(data, assessment, info, type = "data", xykey.cex = 1.4)
```


#### Auxiliary data

```{r, include = FALSE}
ok <- params$compartment %in% c("biota", "sediment")
```

```{asis, eval = !ok}
<br> 
No auxiliary variables currently plotted.
```

```{r auxiliary_data, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
auxiliary_id <- "default"

if (info$compartment %in% "sediment") {
  if (info$determinand %in% c("CD", "PB")) {
    auxiliary_id <- c("value", "concentration", "AL", "LI")
  } else {
    auxiliary_id <- c("value", "concentration", "CORG", "LOIGN")
  }
}

plot.auxiliary(data, info, auxiliary_id = auxiliary_id, xykey.cex = 1.2)
```





#### Assessments (related `r txt_compounds`)

```{r, include = FALSE}
ok <- is_multi
```

```{asis, eval = !ok}
<br>
No related responses assessed.
```

```{r multi_assessment, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.multiassessment(
  assessment_object$data, assessment_object$assessment, info_multi, type = "assessment")
```


#### Data (related `r txt_compounds`)

```{r, include = FALSE}
ok <- is_multi
```

```{asis, eval = !ok}
<br>
No related responses assessed.
```

```{r multi_data, eval = ok, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.multidata(assessment_object$data, info_multi)
```


```{asis, eval = is_ratio}
#### Contaminant ratios
```

```{r ratio_plot, eval = is_ratio, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 7}
plot.ratio(assessment_object$data, info_multi)
```


#### Statistical analysis

<br>

**Trend assessment**


```{asis, eval = !anova_ok & !info$group %in% c("Effects", "Imposex")}
There are too few years, or too many less-than values, for a trend assessment.
<br>
```

```{asis, eval = !anova_ok & info$group %in% "Effects"}
There are too few years for a trend assessment.
<br>
```

```{asis, eval = !anova_ok & info$group %in% "Imposex"}
There is no trend assessment because there are too few years (or occasionally because all the measurements are zero).
<br>
```


```{r, eval = anova_ok & info$detGroup != "Imposex", child = "assessment_platform_trend.Rmd"}
```

```{r, eval = anova_ok & info$detGroup == "Imposex", child = "assessment_platform_trend_imposex.Rmd"}
```

<br>

**Status assessment**

```{r status_ok, include = FALSE}

nyear <- assessment$summary$nyfit
nyear_ok <- !is.null(nyear) && nyear >= 3

status_parametric <- is_AC && nyear_ok

# only need nonparametric analysis if no parametric analysis!

status_nonparametric <- !status_parametric && status_nonparametric

status_ok <- status_parametric || status_nonparametric

if (status_nonparametric) {
  is_below <- assessment$summary[paste0(AC_id, "below")]
  is_below <- unlist(is_below)
  is_below <- is_below %in% "below"
  names(is_below) <- AC_id
  
  status_txt <- mapply(AC_id, is_below, FUN = function(id, below) {
    if (below) {
      "significantly below the threshold value (p < 0.05)"
    } else {
      "not significantly below the threshold value (p > 0.05)"
    }
  })
}

if (!status_ok) {
  is_below <- assessment$summary$clLY < assessment$summary[[AC_id]]

  status_txt <- if (is_below) {
    paste("significantly below the threshold value (p < 0.05).")
  } else {
    paste("not significantly below the threshold value (p > 0.05).")
  }
}


```






```{asis, eval = !status_ok}
The mean `r txt_concentration` was `r status_txt[1]`
<br>
```

```{asis, eval = status_nonparametric && length(AC_id) == 1}
A sign test based on the last five monitoring years shows that the mean `r txt_concentration` was `r status_txt[1]`.
<br>
```

```{asis, eval = status_nonparametric && length(AC_id) == 2}
A sign test based on the last five monitoring years shows that the mean `r txt_concentration` was:  

<ul>
<li>`r status_txt[1]`</li>
<li>`r status_txt[2]`</li>
</ul>
<br>
```

```{asis, eval = status_nonparametric && length(AC_id) == 3}
A sign test based on the last five monitoring years shows that the mean `r txt_concentration` was:  

<ul>
<li>`r status_txt[1]`</li>
<li>`r status_txt[2]`</li>
<li>`r status_txt[3]`</li>
</ul>
<br>
```


```{r, eval = status_parametric & info$detGroup != "Imposex", child = "assessment_platform_status.Rmd"}
```

```{r, eval = status_parametric & info$detGroup == "Imposex", child = "assessment_platform_status_imposex.Rmd"}
```


<br>

```{r interpretation_setup, include = FALSE}
format_p <- function(p) {
  if (p >= 0.0001) {
    p <- round(p, 4)
    paste0("(p = ", format(p, scientific = FALSE, nsmall = 4), ")")
  } else {
    "(p < 0.0001)"
  }
}
```


```{r interpretation_trend, eval = anova_ok, include = FALSE}
wk <- assessment$summary
wk_year <- unique(assessment$data$year)
wk_n <- length(wk_year)
wk_seq <- length(wk_year) == max(wk_year) - min(wk_year) + 1
wk_recent <- info$maxYear - info$recent.trend + 1


change_txt <- function(type = c("whole", "recent")) {
  p <- switch(type, whole = wk$pltrend, wk$prtrend)
  trend <- switch(type, whole = wk$ltrend, wk$rtrend)
  paste0(
    stringr::str_to_sentence(txt_concentrations),
    " at the end of the time series were ", 
    if (p < 0.05) {
      paste("signficantly", if (trend > 0) "higher" else "lower", "than those ") 
    } else {
      "not significantly different from those "
    },
    switch(
      type, 
      whole = "at the start of the time series", 
      paste("in", wk_recent)
    ), 
    " ", format_p(p), ". "
  )
}


if (wk$p_overall >= 0.05) {
  
  trend_description <- paste0(
    "There is no significant temporal trend in the time series ", 
    format_p(wk$p_overall), 
    "."
  )
  
} else if (is.na(wk$p_nonlinear)) {
  
  if (!info$distribution %in% c("lognormal", "normal", "multinomial", "survival")) {
    stop("not coded for ", info$distribution, " distribution")
  }
 
  ltrend_txt1 <- switch(
    info$distribution, 
    lognormal = "log-linear ",
    normal = "", 
    multinomial = "linear logistic ",
    survival = "log-linear"
  )
  
  ltrend_txt2 <- paste0(
    "There is a significant ", 
    ltrend_txt1, 
    "trend in the time series ", 
    format_p(wk$p_linear), 
    ". "
  )

  ltrend_txt3 <- switch(
    info$distribution,
    lognormal = paste0(
      " by an estimated ",
      round(100 * (exp(wk$ltrend / 100) - 1), 1), 
      "% per year over the course of the time series."
    ),
    normal = paste(
      " by an estimated ",
      format(wk$ltrend, digits = 2, nsmall = 0), info$unit_text, 
      "per year over the course of the time series."
    ),
    multinomial = paste(
      " over the course of the time series.",
      if (is_AC && "EAC" %in% AC_id) paste0(
        " Formally, the odds of an individual exceeding the EAC in one year", 
        " relative to the previous year is ",
        format(exp(wk$ltrend), digits = 2, nsmall = 0), 
        "."
      )
    ),
    survival = paste0(
      " by an estimated ",
      round(100 * (exp(wk$ltrend / 100) - 1), 1), 
      "% per year over the course of the time series."
    )
  )


  if (!is_change) {
    
    # not a change point model (currently only imposex)
    
    trend_description <- paste0(
      ltrend_txt2, 
      stringr::str_to_sentence(txt_concentrations),
      " have ", 
      if (wk$ltrend > 0) "increased" else "decreased", 
      ltrend_txt3
    )
    
    if (info$good.status == "high") {
      trend_description <- paste(
        trend_description, 
        "Higher values of",
        info$det_name, 
        "indicate healthier organisms, so the trend suggests a",
        if (wk$ltrend > 0) "improving" else "deteriorating",
        "situation."
      )
    }
    
  } else {
    
    # change point model with linear trend after change
    
    trend_description <- paste0(
      "There is a significant trend in the time series ", 
      format_p(wk$p_overall),  
      " with levels stable until ", 
      modelID, 
      " and then ", 
      if (wk$ltrend > 0) "increasing" else "decreasing", 
      " linear logistically. ", 
      change_txt("whole"), 
      if (assessment$contrasts["whole", "start"] < wk_recent) change_txt("recent")
    )

  }

} else {
  
  if (!is_change) {
  
    # not a change point model (currently only imposex)
  
    trend_description <- paste0(
      "There is a significant trend in the time series ", 
      format_p(wk$p_overall), 
      ". The trend is nonlinear ", 
      format_p(wk$p_nonlinear), 
      " so the shape of the trend must be assessed visually. ", 
      change_txt("whole"), 
      if (assessment$contrasts["whole", "start"] < wk_recent) change_txt("recent")
    )

  } else if (grepl("smooth", assessment$method)) {
    
    # change point model with nonlinear trend after change
    
    trend_description <- paste0(
      "There is a significant temporal trend in the time series ", 
      format_p(wk$p_overall), 
      ". The trend is nonlinear ", 
      format_p(wk$p_nonlinear), 
      " with levels stable until ", 
      modelID, 
      " followed by a nonlinear pattern of change which must be assessed visually. ", 
      change_txt("whole"), 
      if (assessment$contrasts["whole", "start"] < wk_recent) change_txt("recent")
    )
    
  } 
    

  if (info$good.status %in% "high") {
    trend_description <- paste(
      trend_description, 
      "Note that ", 
      if (wk$ltrend > 0) "higher" else "lower",
      "values of", 
      info$det_name, 
      "indicate",
      if (wk$ltrend > 0) "healthier"else "less healthy",
      "organisms."
    )
  }

} 
```


```{r interpretation_status, include = FALSE}

if (status_parametric && info$detGroup != "Imposex") {

  AC_p <- rv[[p_txt]]
  names(AC_p) <- row.names(rv)
  
  p_value <- sapply(AC_p, format_p)
  is_below <- sapply(AC_p, function(x) x < 0.05)

} 

if (status_parametric && info$detGroup == "Imposex") {

  p_value <- rv[["Pr(>Ref)"]]
  names(p_value) <- row.names(rv)
  
  p_value <- ifelse(p_value == "<0.05", "(p < 0.05)", "(p > 0.05)")
  
  is_below <- p_value == "(p < 0.05)"
} 


if (!status_ok) {

  p_value <- ifelse(is_below, "(p < 0.05)", "(p > 0.05)")  

}  

below_txt <- switch(
  info$good.status, 
  low = "below", 
  high = "above"
)

above_txt <- switch(
  info$good.status,
  low = "above", 
  high = "below"
)

if (is_below) {
  
  status_txt1 <- paste0(
    "significantly ", below_txt, " the threshold value ", p_value, "."
  )
  
  status_txt2 <- "were therefore unlikely to have harmed marine life."

  if (info$group %in% c("Effects", "Imposex")) {
    status_txt2 <- "were therefore unlikely to indicate harm."
  }
    
} else {
  
  status_txt1 <- paste0(
    "not significantly ", below_txt, " the threshold value ", p_value, "."
  )

  status_txt2 <- "could therefore have harmed marine life."
  
  if (info$group %in% c("Effects", "Imposex")) {
    status_txt2 <- "could therefore indicate harm."
  }

}  


status_description <- paste(
  stringr::str_to_sentence(txt_concentrations),
  "at the end of the time series were",
  status_txt1,
  stringr::str_to_sentence(txt_concentrations),
  status_txt2
)
```



```{asis}
**Interpretation**

`r if (anova_ok) trend_description`

`r status_description`

<br>

```




```{asis, include = FALSE, eval = FALSE}
The lowest detectable annual increase in the time series (two-sided test, power = 80%, size = 5%) is `r wk$dtrend_1`%. `r if (!wk_seq) paste0("Had the data been collected sequentially (rather than with gaps between years) the lowest detectable annual increase would be ", wk$dtrend_2, "%.")`  The lowest detectable annual increase with 10 years of `r if (!wk_seq) "sequential"` monitoring is `r wk$dtrend_3`%.

Given the variability in the data, `r wk$dyear` years of `r if (!wk_seq) "sequential"` monitoring would be required to detect an annual increase of 10% with 80% power (two-sided test, size = 5%).

The power of the time series to detect an annual increase of 10% is `r wk$dpower_1`% (two-sided test, size = 5%).  `r if (!wk_seq) paste0("With sequential monitoring, the power would be ", wk$dpower_2, "%")`.  With 10 years of `r if (!wk_seq) "sequential"` monitoring, the power would be `r wk$dpower_3`%.
```


```{r tidy_up, include = FALSE}
rm(list = intersect(
  objects(),
  c("above_txt", "AC_id", "AC_ok", "AC_p", "AC_type", 
    "anova", "anova_ok", "args.list", "assessment", "assessment_object", 
    "auxiliary_id", 
    "BAC_id", "below_txt", "change_txt", "coefficients", "contrasts", "data", 
    "determinands", "diff_txt", "dups", "EAC_id", "extra.text", "format_p", 
    "good_status", "group_id", "HQS_id", 
    "id", "info", "info_multi", "intro_AMAP", "intro_txt",
    "is_AC", "is_BAC", "is_below", "is_EAC", "is_HQS", "is_normalised", "is_pred", 
    "is_ratio", "is_lognormal", "ltrend_txt",
    "nyear", "nyear_ok", "ok", "p_txt", "p_value", "page_title", "pred", "rv",
    "species_id",
    "status_description", "status_ok", "status_parametric", "status_nonparametric",
    "status_txt", "status_txt1", "status_txt2", "status_txt3", "status_txt4", 
    "trend_description", 
    "txt_AC", "txt_compounds", "txt_concentration", "txt_concentrations", 
    "txt_measurements", "txt_related1", "txt_related2", "txt_trend1", "txt_trend2", 
    "txt_trend3",
    "var_id", "wk", "wk_n", "wk_recent", "wk_seq", "wk_year"
  )
))
```


